Revisiting the Relationship between Short-Chain Fatty Acids, the Microbiota, and Colorectal Tumors

### 2019-03-29 submission/manuscript.tex
Edit manuscript before sending to co-authors

The gut microbiome is thought to have a role in the development of
colorectal cancer by protecting against and exacerbating inflammation.
Short chain fatty acids (SCFAs), including butyrate, have been shown to
have anti-inflammatory properties and are produced in large quantities
by colonic bacteria. We assessed whether there was an association
between fecal SCFA concentrations and the presence of colonic adenomas
or carcinomas in a cohort of individuals that was previously subjected
to 16S rRNA gene and metagenomic shotgun sequencing. We measured the
fecal concentrations of acetate, propionate, isobutyrate, and butyrate
within the cohort and found that there were no meaningful differences in
their concentration and disease status. When we incorporated these
concentrations into Random Forest classification models trained to
differentiate between people with normal colons and those with adenomas
or carcinomas, we found that they did not significantly improve the
ability of 16S rRNA gene or metagenomic gene sequence-based models to
classify individuals. Finally, we generated Random Forest regression
models trained to predict the concentration of each SCFA based on 16S
rRNA gene or metagenomic gene sequence data from the same samples. These
models performed poorly and were only able to explain less than 10\% of
the observed variation in the SCFA concentrations. These results support
the broader epidemiological data that questions the value of fiber
consumption for reducing the risks of colorectal cancer. Although it is
likely that bacterial metabolites may serve as biomarkers to detect
adenomas or carcinomas, fecal SCFA concentrations have limited value.

@@

SCFAs have have anti-inflammatory and anti-proliferative activities (9).
Manipulation of SCFAs in mouse models of colorectal cancer by direct
supplementation or feeding of fiber caused an overall reduction in tumor
burden (10). These results suggest that supplementation with substrates
that bacteria can ferment to produce SCFAs may confer beneficial effects
against colorectal cancer. Regardless, there is a lack of evidence that
increasing SCFA concentrations can protect against colorectal cancer in
humans. Based on similar observations, many microbiome studies use the
concentrations of SCFAs and the presence of 16S rRNA gene sequences from
organisms and the genes involved in producing them as a biomarker of a
healthy microbiota (11, 12). Case-control studies that have investigated
SCFA concentrations in colorectal cancer found that patients with
carcinomas had lower concentrations of SCFAs versus patients with
adenomas or individuals without colon tumors (13). Although this would
argue that increasing SCFA concentrations could be protective against
tumorigenesis, in randomized controlled trials fiber supplementation has
been inconsistently associated with protection against tumor formation
and recurrence (14, 15). These findings temper enthusiasm for treatments
that aim to use SCFAs as biomarkers or protection against tumorigenesis.

\textbf{SCFA concentrations do not meaningfully vary with diagnosis or
treatment.} To quantify the associations between colorectal cancer, the
microbiome, and SCFAs, we quantified the concentration of acetate,
propionate, isobutyrate, and butyrate in feces of previously
characterized individuals with normal colons (N=172) and those with
colonic adenomas (N=198) or carcinomas (N=120) (16). The only SCFA that
had a significantly different concentration across the diagnoses was
isobutyrate (P=0.0091; Figure 1A). Interestingly, the median
concentration of isobutyrate was 3.30 mmol/kg in people with normal
colons and it was 3.00 and 3.84 mmol/kg in people with colonic adenomas
or carcinomas, respectively. The difference in isobutyrate concentration
between people with adenomas and carcinomas was significantly different
(P=0.0065); however, the differences between people with normal colons
and those with adenomas or carcinomas was not significant (P=0.19 and
P=0.11). Among the subjects with adenomas and carcinomas, a subset
(N\textsubscript{adenoma}=41, N\textsubscript{carcinoma}=26) were
treated and sampled a year later (17). The only SCFA that changed
following treatment was isobutyrate, which decreased by 0.99 mmol/kg
(P=0.002; Figure 1B). For both the pre-treatment cross-sectional data
and the pre/post treatment data, we pooled the SCFA concentrations on a
per molecule of carbon basis and again failed to see any significant
differences (P\textgreater{}0.15). The low concentration of isobutyrate
relative to the other SCFAs, inconsistent concentrations, and unexpected
decrease in concentration with treatment makes it difficult to ascribe
much biological relevance to this observation.

\textbf{Combining SCFA and microbiome data does not improve the ability
to diagnose individual as having adenomas or carcinomas.} We previously
found that binning 16S rRNA gene sequence data into operational
taxonomic units based on 97\% similarity or into genera enabled us to
classify individuals as having adenomas or carcinomas using Random
Forest machine learning models (8, 16). We repeated that analysis but
added the concentration of the SCFAs as possible features to train the
models (Figure S1). Models trained using SCFAs to classify individuals
as having adenomas or carcinomas rather than normal colons had median
areas under the receiver operator characteristic curve (AUROC) that were
significantly greater than 0.5 (P\textsubscript{adenoma}\textless{}0.001
and P\textsubscript{carcinoma}\textless{}0.001); however, the AUROC
values to detect the presence of adenomas or carcinomas were only 0.54
and 0.55, respectively (Figure 2A). When we trained the models with the
SCFAs concentrations and operational taxonomic unit (OTU) or genus-level
relative abundances the AUROC values were not significantly different
from the models trained without the SCFA concentrations
(P\textgreater{}0.21; Figure 2A). These data demonstrate that knowledge
of the SCFA profile from a patient's fecal sample does not improve the
ability to diagnose a colonic lesion.

\textbf{Knowledge of microbial community structure does not predict SCFA
concentrations.} We next asked whether the fecal community structure was
predictive of fecal SCFA concentrations, regardless of a person's
diagnosis. We trained Random Forest regression models using 16S rRNA
gene sequence data binned into OTUs and genera to predict the
concentration of the SCFAs (Figure S2). The largest R\textsuperscript{2}
between the observed SCFA concentrations and the modeled concentrations
was 0.14 observed when using genus data to predict butyrate
concentrations (Figure 2B). We also used a smaller dataset of shotgun
metagenomic sequencing data generated from a subset of our cohort
(N\textsubscript{normal}=27, N\textsubscript{adenoma}=25, and
N\textsubscript{cancer}=26) (18). We binned genes extracted from the
assembled metagenomes into operational protein families (OPFs) or KEGG
categories and trained Random Forest regression models using metagenomic
to predict the concentration of the SCFAs (Figure S2). Similar to the
analysis using 16S rRNA gene sequence data, the metagenomic data was not
predictive of SCFA concentration; the largest amount of variation that
the models could explain was 0.055, which was observed when using KEGG
data to predict propionate concentrations (Figure 2B). Because of the
limited number of samples that we were able to generate metagenomic
sequence data from, we used our 16S rRNA gene sequence data to impute
metagenomes that were binned into metabolic pathways or KEGG categories
using PICRUSt (Figure S2). SCFA concentrations could not be predicted
based on the imputed metagenomic data. The largest amount of variation
that the models could explain was 0.097 observed when using KEGG data to
predict butyrate concentrations (Figure 2B). The inability to model SCFA
concentrations from microbiome data indicates that the knowledge of the
abundance of organisms and their genes was insufficient to predict SCFA
concentrations.

\textbf{Conclusion.} Our data indicate that fecal SCFA concentrations
are not associated with the presence of adenomas or carcinomas and that
they provide weak predictive power to improve the ability to diagnose
someone with one of these lesions. Furthermore, knowledge of the
taxonomic and genetic structure of gut microbiota was not predictive
SCFA concentrations. These results complement existing literature that
suggest that fiber consumption and the production of SCFAs are unable to
prevent the risk of developing colonic tumors. It is important to note
that our analysis concerned fecal SCFA concentrations and microbiome
characterization and that observations along the mucosa near the site of
lesions may provide a stronger association. Regardless, given the
growing literature in this area, it is unlikely that SCFAs are the
primary mechanism that limits tumorigenesis. This may be a cautionary
result to temper enthusiasm for SCFAs as a biomarker of gut health more
generally. Going forward it is critical to develop additional hypotheses
for how the microbiome and host interact to drive tumorigenesis to
better understand the disease process and identify biomarkers that will
allow early detection of tumors.

@@

\textbf{16S rRNA gene sequence data analysis.} Sequence data from Baxter
et al.~(16) and Sze et al.~(17) were obtained from the Sequence Read
Archive (studies SRP062005 and SRP096978) and reprocessed using using
mothur v.1.42 (20). The original studies generated sequence data from V4
region of the 16S rRNA gene using paired 250 nt reads on an Illumina
MiSeq sequencer. The resulting sequence data were assembled into
contigs, screened to remove low quality contigs and chimeras. The
curated sequences were then clustered into OTUs at a 97\% similarity
threshold and assigned to the closest possible genus with an 80\%
confidence threshold trained on the reference collection from the
Ribosomal Database Project (v.16). We used PICRUSt (v.2.1.0-b) with the
recommended standard operating protocol to generate imputed metagenomes
based on the expected metabolic pathways, KEGG categories, and enzyme
commission numbers (21).

\textbf{Metagenomic DNA sequence analysis.} A subset of the samples from
the samples described by Baxter et al.~(16) were used to generate
metagenomic sequence data (N\textsubscript{normal}=27,
N\textsubscript{adenoma}=25, and N\textsubscript{cancer}=26). These data
were generated by Hannigan et al.~(18) and deposited into the Sequence
Read Archive (study SRP108915). Fecal DNA was subjected to shotgun
sequencing on an Illumina HiSeq using 125 bp paired end reads. The
archived sequences were already quality filtered and aligned to the
human genome to remove contaminating sequence data. We downloaded the
sequences and assembled them into contigs using MEGAHIT (22), which were
used to identify open reading frames (ORFs) using Prodigal (23). We
determined the abundance of each ORF by mapping the raw reads back to
the ORFs using Diamond (24). We clustered the ORFs into operational
protein families (OPFs) in which the clustered ORFs were more than 40\%
identical to each other using mmseq2 (25). We also used mmseq2 to map
the ORFs to the KEGG database and clustered the ORFs according to which
category the ORFs mapped.

\textbf{Random Forest models.} The classification models were built to
predict lesion type from microbiome information with or without SCFA
concentrations. The regression models were built to predict the SCFA
concentrations of acetate, butyrate, and propionate from microbiome
information. For classification and regression models, we pre-processed
the features by scaling them to vary between zero and one. Features with
no variance in the training set were removed from both the training and
testing sets. We randomly split the data into training and test sets so
that the training set consisted of 80\% of the full dataset while the
test set was composed of the remaining data. The training set was used
for hyperparameter selection and training the model, and the test set
was used for evaluating prediction performance. For each model, the best
performing hyperparameter, mtry, was selected in an internal five-fold
cross-validation of the training set with 100 randomizations. Six values
of mtry were tested and the value that provided the largest AUROC or
R\textsuperscript{2} was selected. We trained the random forest model
using the selected mtry value and predicted the held-out test set. The
data-split, hyperparameter selection, training and testing steps were
repeated 100 times to get a reliable and robust reading of model
prediction performance. We used AUROC and R\textsuperscript{2} as the
prediction performance metric for classification and regression models,
respectively. We used randomForest package implemented to the caret
package (version 4.6-14) in R statistical software (version 6.0-81) for
our models.

@@

\leavevmode\hypertarget{ref-Ohigashi2013}{}%
13. \textbf{Ohigashi S}, \textbf{Sudo K}, \textbf{Kobayashi D},
\textbf{Takahashi O}, \textbf{Takahashi T}, \textbf{Asahara T},
\textbf{Nomoto K}, \textbf{Onodera H}. 2013. Changes of the intestinal
microbiota, short chain fatty acids, and fecal pH in patients with
colorectal cancer. Digestive Diseases and Sciences
\textbf{58}:1717--1726.
doi:\href{https://doi.org/10.1007/s10620-012-2526-4}{10.1007/s10620-012-2526-4}.

@@

\textbf{Figure 1. SCFA concentrations did not vary meaningfully with
diagnosis of colonic lesions or with treatment for adenomas or
carcinomas.} (A) We measured the concentration of fecal SCFAs from
individuals with normal colons (N=172) or those with adenoma (N=198) or
carcinomas (N=120). (B) A subset of individuals diagnosed with adenomas
(N=41) or carcinomas (N=26) who underwent treatment were resampled a
year after the initial sampling; one extreme propionate value (124.4
mmol/kg) was included in the adenoma analysis but censored from the
visualization for clarity.

@@

\textbf{Figure S1. Comparison of training and testing results for
classification models shows that the models are robust and are not
overfit.} Random Forest classification models were generated to
differentiate between subjects with normal colons and those with
adenomas or carcinomas using 16S rRNA gene sequence data that were
clustered into genera or OTUs with and without including the four SCFAs
as additional features. Random Forest classification models were
generated by partitioning the samples into a training set with 80\% of
the data and a testing set with the remaining samples. The training data
was used to fit the mtry hyperparameter using 100 five-fold cross
validations for six values of mtry. The mtry with the largest AUROC was
selected to train the model, which was then tested on the held out 20\%
of the samples.

@@

\textbf{Figure S2. Comparison of training and testing results for
regression models shows that the models are robust and are not overfit.}
Random Forest regression models were generated to predict the
concentration of each SCFA using each subjects' microbiome data
generated using 16S rRNA gene sequence and metagenomic sequence data.
These regression models were generated by partitioning the samples into
a training set with 80\% of the data and a testing set with the
remaining samples. The training data was used to fit the mtry
hyperparameter using 100 five-fold cross validations for six values of
mtry. The mtry with the largest R\textsuperscript{2} was selected to
train the model, which was then tested on the held out 20\% of the
samples.

>>>>

The gut microbiome is thought to have a role in the development of
colorectal cancer by protecting against and exacerbating inflammation.
Short chain fatty acids (SCFAs), including butyrate, have been shown to
have anti-inflammatory properties and are produced in large quantities
by colonic bacteria. We assessed whether there was an association
between fecal SCFA concentrations and the presence of colonic adenomas
or carcinomas in a cohort of individuals that was previously
individualed to 16S rRNA gene and metagenomic shotgun sequencing. We
measured the fecal concentrations of acetate, propionate, isobutyrate,
and butyrate within the cohort and found that there were no meaningful
differences in their concentration and disease status. When we
incorporated these concentrations into Random Forest classification
models trained to differentiate between people with normal colons and
those with adenomas or carcinomas, we found that they did not
significantly improve the ability of 16S rRNA gene or metagenomic gene
sequence-based models to classify individuals. Finally, we generated
Random Forest regression models trained to predict the concentration of
each SCFA based on 16S rRNA gene or metagenomic gene sequence data from
the same samples. These models performed poorly and were able to explain
at most 14\% of the observed variation in the SCFA concentrations. These
results support the broader epidemiological data that questions the
value of fiber consumption for reducing the risks of colorectal cancer.
Although it is likely that bacterial metabolites may serve as biomarkers
to detect adenomas or carcinomas, fecal SCFA concentrations have limited
value.

@@

SCFAs have anti-inflammatory and anti-proliferative activities (9).
Direct supplementation of SCFAs or feeding of fiber caused an overall
reduction in tumor burdenin mouse models of colorectal cancer (10).
These results suggest that supplementation with substrates that bacteria
can ferment to produce SCFAs may confer beneficial effects against
colorectal cancer. Based on similar observations, many microbiome
studies use the concentrations of SCFAs and the presence of 16S rRNA
gene sequences from organisms and the genes involved in producing them
as a biomarker of a healthy microbiota (11, 12). Regardless, there is a
lack of evidence that increasing SCFA concentrations can protect against
colorectal cancer in humans. Case-control studies that have investigated
a possible associations between SCFAs and colon tumor status have been
plagued by relatively small numbers of subjects, but have reported
increased total and relative fecal acetate levels and decreased relative
fecal butyrate concentrations in subjects with colonic lesions (13).
Based on these limited results, one would hypothesize increasing SCFA
concentrations could limit tumorigenesis, in randomized controlled
trials fiber supplementation has been inconsistently associated with
protection against tumor formation and recurrence (14, 15). These
findings temper enthusiasm for treatments that aim to use SCFAs as
biomarkers or protection against tumorigenesis.

\textbf{SCFA concentrations do not meaningfully vary with diagnosis or
treatment.} To quantify the associations between colorectal cancer, the
microbiome, and SCFAs, we quantified the concentration of acetate,
propionate, isobutyrate, and butyrate in feces of previously
characterized individuals with normal colons (N=172) and those with
colonic adenomas (N=198) or carcinomas (N=120) (16). The only SCFA that
had a significantly different concentration across the diagnoses was
isobutyrate (P=0.0091; Figure 1A). Interestingly, the median
concentration of isobutyrate was 3.30 mmol/kg of feces in people with
normal colons and it was 3.00 and 3.84 mmol/kg of feces in people with
colonic adenomas or carcinomas, respectively. The difference in
isobutyrate concentration between people with adenomas and carcinomas
was significantly different (P=0.0065); however, the differences between
people with normal colons and those with adenomas or carcinomas was not
significant (P=0.19 and P=0.11). Among the individuals with adenomas and
carcinomas, a subset (N\textsubscript{adenoma}=41,
N\textsubscript{carcinoma}=26) were treated and sampled a year later
(17). The only SCFA that changed following treatment was isobutyrate,
which decreased by 0.99 mmol/kg of feces (P=0.002; Figure 1B). For both
the pre-treatment cross-sectional data and the pre/post treatment data,
we also failed to detect any significant differences in the relative
concentrations of any SCFAs other than isobutyrate (0.17). Finally, we
pooled the SCFA concentrations on a total and per molecule of carbon
basis and again failed to observe any significant differences
(P\textgreater{}0.15). The low concentration of isobutyrate relative to
the other SCFAs, inconsistent concentrations, and unexpected decrease in
concentration with treatment makes it difficult to ascribe much
biological relevance to the isobutyrate observations.

\textbf{Combining SCFA and microbiome data does not improve the ability
to diagnose individual as having adenomas or carcinomas.} We previously
found that binning 16S rRNA gene sequence data into operational
taxonomic units (OTUs) based on 97\% similarity or into genera enabled
us to classify individuals as having adenomas or carcinomas using Random
Forest machine learning models (8, 16). We repeated that analysis but
added the concentration of the SCFAs as possible features to train the
models (Figure S1). Models trained using SCFAs to classify individuals
as having adenomas or carcinomas rather than normal colons had median
areas under the receiver operator characteristic curve (AUROC) that were
significantly greater than 0.5 (P\textsubscript{adenoma}\textless{}0.001
and P\textsubscript{carcinoma}\textless{}0.001); however, the AUROC
values to detect the presence of adenomas or carcinomas were only 0.54
and 0.55, respectively (Figure 2A). When we trained the models with the
SCFAs concentrations and OTU or genus-level relative abundances the
AUROC values were not significantly different from the same models
trained without the SCFA concentrations (P\textgreater{}0.21; Figure
2A). These data demonstrate that knowledge of the SCFA profile from a
patient's fecal sample did not improve the ability to diagnose a colonic
lesion.

\textbf{Knowledge of microbial community structure does not predict SCFA
concentrations.} We next asked whether the fecal community structure was
predictive of fecal SCFA concentrations, regardless of a person's
diagnosis. We trained Random Forest regression models using 16S rRNA
gene sequence data binned into OTUs and genera to predict the
concentration of the SCFAs (Figure S2). The largest R\textsuperscript{2}
between the observed SCFA concentrations and the modeled concentrations
was 0.14, which was observed when using genus data to predict butyrate
concentrations (Figure 2B). We also used a smaller dataset of shotgun
metagenomic sequencing data generated from a subset of our cohort
(N\textsubscript{normal}=27, N\textsubscript{adenoma}=25, and
N\textsubscript{cancer}=26) (18). We binned genes extracted from the
assembled metagenomes into operational protein families (OPFs) or KEGG
categories and trained Random Forest regression models using metagenomic
to predict the concentration of the SCFAs (Figure S2). Similar to the
analysis using 16S rRNA gene sequence data, the metagenomic data was not
predictive of SCFA concentration; the largest amount of variation that
the models could explain was 0.055, which was observed when using KEGG
data to predict propionate concentrations (Figure 2B). Because of the
limited number of samples that we were able to generate metagenomic
sequence data from, we used our 16S rRNA gene sequence data to impute
metagenomes that were binned into metabolic pathways or KEGG categories
using PICRUSt (Figure S2). SCFA concentrations could not be predicted
based on the imputed metagenomic data. The largest amount of variation
that the models could explain was 0.097, which was observed when using
KEGG data to predict butyrate concentrations (Figure 2B). The inability
to model SCFA concentrations from microbiome data indicates that the
knowledge of the abundance of organisms and their genes was insufficient
to predict SCFA concentrations.

\textbf{Conclusion.} Our data indicate that fecal SCFA concentrations
are not associated with the presence of adenomas or carcinomas and that
they provide weak predictive power to improve the ability to diagnose
someone with one of these lesions. Furthermore, knowledge of the
taxonomic and genetic structure of gut microbiota was not predictive
SCFA concentrations. These results complement existing literature that
suggest that fiber consumption and the production of SCFAs are unable to
prevent the risk of developing colonic tumors. It is important to note
that our analysis concerned fecal SCFA concentrations and microbiome
characterization and that observations along the mucosa near the site of
lesions may provide a stronger association. Regardless, given the
growing literature in this area, it is unlikely that SCFAs are the
primary mechanism that limits tumorigenesis. This may be a cautionary
result to temper enthusiasm for SCFAs as a biomarker of gut health more
generally. Going forward it is critical to develop additional hypotheses
for how the microbiome and host interact to drive tumorigenesis so that
we can better understand tumorigenesis and identify biomarkers that will
allow early detection of lesions.

The authors thank the Great Lakes-New England Early Detection Research
Network for providing the fecal samples that were used in this study. We
would thank the University of Michigan Center for Microbial Systems for
enabling our short-chain fatty acid analysis. Support for MAS came from
the Canadian Institute of Health Research and the National Institutes of
Health (UL1TR002240). This work was also supported by the National
Institutes of Health (P30DK034933 and R01CA215574).

@@

\textbf{16S rRNA gene sequence data analysis.} Sequence data from Baxter
et al.~(16) and Sze et al.~(17) were obtained from the Sequence Read
Archive (studies SRP062005 and SRP096978) and reprocessed using using
mothur v.1.42 (20). The original studies generated sequence data from V4
region of the 16S rRNA gene using paired 250 nt reads on an Illumina
MiSeq sequencer. The resulting sequence data were assembled into contigs
and screened to remove low quality contigs and chimeras. The curated
sequences were then clustered into OTUs at a 97\% similarity threshold
and assigned to the closest possible genus with an 80\% confidence
threshold trained on the reference collection from the Ribosomal
Database Project (v.16). We used PICRUSt (v.2.1.0-b) with the
recommended standard operating protocol to generate imputed metagenomes
based on the expected metabolic pathways and KEGG categories (21).

\textbf{Metagenomic DNA sequence analysis.} A subset of the samples from
the samples described by Baxter et al.~(16) were used to generate
metagenomic sequence data (N\textsubscript{normal}=27,
N\textsubscript{adenoma}=25, and N\textsubscript{cancer}=26). These data
were generated by Hannigan et al.~(18) and deposited into the Sequence
Read Archive (study SRP108915). Fecal DNA was individualed to shotgun
sequencing on an Illumina HiSeq using 125 bp paired end reads. The
archived sequences were already quality filtered and aligned to the
human genome to remove contaminating sequence data. We downloaded the
sequences and assembled them into contigs using MEGAHIT (22), which were
used to identify open reading frames (ORFs) using Prodigal (23). We
determined the abundance of each ORF by mapping the raw reads back to
the ORFs using Diamond (24). We clustered the ORFs into operational
protein families (OPFs) in which the clustered ORFs were more than 40\%
identical to each other using mmseq2 (25). We also used mmseq2 to map
the ORFs to the KEGG database and clustered the ORFs according to which
category the ORFs mapped.

\textbf{Random Forest models.} The classification models were built to
predict lesion type from microbiome information with or without SCFA
concentrations. The regression models were built to predict the SCFA
concentrations of acetate, butyrate, and propionate from microbiome
information. For classification and regression models, we pre-processed
the features by scaling them to vary between zero and one. Features with
no variance in the training set were removed from both the training and
testing sets. We randomly split the data into training and test sets so
that the training set consisted of 80\% of the full dataset while the
test set was composed of the remaining data. The training set was used
for hyperparameter selection and training the model and the test set was
used for evaluating prediction performance. For each model, the best
performing hyperparameter, mtry, was selected in an internal five-fold
cross-validation of the training set with 100 randomizations. Six values
of mtry were tested and the value that provided the largest AUROC or
R\textsuperscript{2} was selected. We trained the random forest model
using the selected mtry value and predicted the held-out test set. The
data-split, hyperparameter selection, training and testing steps were
repeated 100 times to get a reliable and robust reading of model
prediction performance. We used AUROC and R\textsuperscript{2} as the
prediction performance metric for classification and regression models,
respectively. We used randomForest package implemented to the caret
package (version 4.6-14) in R statistical software (version 6.0-81) for
our models.

@@

\leavevmode\hypertarget{ref-Weaver1988}{}%
13. \textbf{Weaver GA}, \textbf{Krause JA}, \textbf{Miller TL},
\textbf{Wolin MJ}. 1988. Short chain fatty acid distributions of enema
samples from a sigmoidoscopy population: An association of high acetate
and low butyrate ratios with adenomatous polyps and colon cancer. Gut
\textbf{29}:1539--1543.
doi:\href{https://doi.org/10.1136/gut.29.11.1539}{10.1136/gut.29.11.1539}.

@@

\textbf{Figure 1. SCFA concentrations did not vary meaningfully with
diagnosis of colonic lesions or with treatment for adenomas or
carcinomas.} (A) The concentration of fecal SCFAs from individuals with
normal colons (N=172) or those with adenoma (N=198) or carcinomas
(N=120). (B) A subset of individuals diagnosed with adenomas (N=41) or
carcinomas (N=26) who underwent treatment were resampled a year after
the initial sampling; one extreme propionate value (124.4 mmol/kg) was
included in the adenoma analysis but censored from the visualization for
clarity.

@@

\textbf{Figure S1. Comparison of training and testing results for
classification models shows that the models are robust and are not
overfit.} Random Forest classification models were generated to
differentiate between individuals with normal colons and those with
adenomas or carcinomas using 16S rRNA gene sequence data that were
clustered into genera or OTUs with and without including the four SCFAs
as additional features. Random Forest classification models were
generated by partitioning the samples into a training set with 80\% of
the data and a testing set with the remaining samples for 100
randomizations.

@@

\textbf{Figure S2. Comparison of training and testing results for
regression models shows that the models are robust and are not overfit.}
Random Forest regression models were generated to predict the
concentration of each SCFA using each individuals' microbiome data
generated using 16S rRNA gene sequence and metagenomic sequence data.
These regression models were generated by partitioning the samples into
a training set with 80\% of the data and a testing set with the
remaining samples for 100 randomizations.
